---
title: "3. Resumo"
---

```{r}
source("_setup.R")
df <- out("out_03_studies")
```

::: {.panel-tabset group="lang"}
### Português
O [`03_ai_summary.R`](`r gh("03_ai_summary.R")`){target="_blank"} segue o mesmo caminho da etapa anterior, mas pede um resumo de três frases sobre o estudo.

### English
[`03_ai_summary.R`](`r gh("03_ai_summary.R")`){target="_blank"} follows the same path as the previous step, but asks for a three-sentence summary of the trial.
:::

## O *prompt*

::: {.panel-tabset group="lang"}
### Português
O resumo deve focar nos tratamentos, nos desfechos e na população do estudo. Um ponto importante é pedir que o modelo não "resuma" os nomes de doenças e medicamentos: eles devem aparecer exatamente como no estudo, porque é isso que o médico procura.

### English
The summary should focus on treatments, outcomes and the study population. One important point is asking the model not to "summarize" disease and drug names: they must appear exactly as in the trial, because that's what physicians look for.
:::

```r
studySummaryPrompt <- "You will receive a study's title and description, and your task is to generate a concise, flowing summary of the study. The paragraph must:

- Be written in both English and Portuguese.
- Contain 3 sentences and cover most key aspects of the study methodology and outcomes. Focusing in the study treatments and outcomes studied.
- Explain the studies target population.
- Do not summarise cancer conditions or medicamentions names. Keep them explicitly stated in the criteria.
- Return the result in JSON format with two variables: 'study_ai_summary_english' and 'study_ai_summary_portuguese'.

The output JSON should look like this:

{{
'study_ai_summary_english' : 'string'
'study_ai_summary_portuguese' : 'string'
}}

The study title is: {var_1}
The study description is: {var_2}"
```

## Aplicando a cada estudo

::: {.panel-tabset group="lang"}
### Português
Agora o script lê o `out_02_studies.json`, que já tem os títulos da etapa anterior. As colunas novas são acrescentadas a ele.

### English
Now the script reads `out_02_studies.json`, which already has the titles from the previous step. The new columns are added to it.
:::

```r
df_studies <- fromJSON('data/out_02_studies.json')

df_studies %>%
      split(1:nrow(.)) %>%
      pblapply(function(x) chatLLM(x, intruction_prompt = studySummaryPrompt,
                                   var_1 = x[['study_official_title']],
                                   var_2 = x[['study_brief_summary']])) %>%
      do.call(dplyr::bind_rows, .) %>%
      toJSON(pretty = TRUE) %>%
      write(file = 'data/out_03_studies.json')
```

## Resultado

::: {.panel-tabset group="lang"}
### Português
```{r}
df %>%
  transmute(study_nct_id, study_ai_summary_portuguese = clip(study_ai_summary_portuguese, 400)) %>%
  head(3) %>%
  kable()
```

### English
```{r}
df %>%
  transmute(study_nct_id, study_ai_summary_english = clip(study_ai_summary_english, 400)) %>%
  head(3) %>%
  kable()
```
:::
