---
title: "4. Paciente ideal"
---

```{r}
source("_setup.R")
df <- out("out_04_studies")
```

::: {.panel-tabset group="lang"}
### Português
Os critérios de elegibilidade são a parte mais difícil de ler em um estudo clínico: listas longas de inclusão e exclusão, muitas vezes contraditórias à primeira vista. O [`04_ai_ideal_patient.R`](`r gh("04_ai_ideal_patient.R")`){target="_blank"} transforma esses critérios em um parágrafo que descreve o paciente que se encaixaria no estudo.

### English
Eligibility criteria are the hardest part of a clinical trial to read: long inclusion and exclusion lists, often seemingly contradictory at first. [`04_ai_ideal_patient.R`](`r gh("04_ai_ideal_patient.R")`){target="_blank"} turns these criteria into a paragraph describing the patient who would fit the trial.
:::

## O *prompt*

::: {.panel-tabset group="lang"}
### Português
Este é o *prompt* mais longo até aqui. Antes de escrever, o modelo é orientado a organizar os critérios por tema (diagnóstico, tratamentos prévios, estado funcional...) e cruzar inclusão com exclusão. Assim, um critério como "carcinoma renal de células não claras" combinado com a exclusão do subtipo papilífero vira uma única frase precisa.

O texto final deve ter cinco frases, começar direto pelas características do paciente e manter os termos médicos exatos. Exames laboratoriais e exigências de contracepção ficam de fora, a não ser que sejam essenciais.

### English
This is the longest prompt so far. Before writing, the model is told to organize the criteria by topic (diagnosis, prior treatments, performance status...) and cross-check inclusion against exclusion. That way, a criterion like "non-clear cell renal carcinoma" combined with excluding the papillary subtype becomes a single precise sentence.

The final text should have five sentences, start directly with the patient's characteristics and keep the exact medical terms. Lab tests and contraception requirements are left out unless essential.
:::

```r
idealPatientPrompt <- "# Study Patient Profile Generator
## Task
You will receive a clinical studys eligibility criteria and generate a concise, flowing paragraph describing the ideal patient profile for this study in both English and Portuguese.

## Chain-of-Thought Process (Do Not Display in Output)
Before generating the final paragraphs, follow these analytical steps:

1. First, organize all eligibility criteria into categorical groups:
   - Primary diagnosis (disease type, stage, classification)
   - Prior treatments (medications, procedures, therapies)
   - Performance status requirements
   - Exclusion criteria
   # ...
2. Identify and reconcile any complementary or contradictory criteria
3. Prioritize the most clinically relevant aspects
4. Draft and refine 5 comprehensive sentences that capture the essence of the eligible patient

## Requirements for the Patient Profile
# ... (12 regras, ver o script)

{{
  \"ideal_patient_english\": \"Your 5-sentence English paragraph here\",
  \"ideal_patient_portuguese\": \"Your 5-sentence Portuguese paragraph here\"
}}

The study eligibility criteria are: {var_1}."
```

## Aplicando a cada estudo

::: {.panel-tabset group="lang"}
### Português
Aqui só uma variável é enviada: os critérios de elegibilidade completos.

### English
Only one variable is sent here: the full eligibility criteria.
:::

```r
df_studies <- fromJSON('data/out_03_studies.json')

df_studies %>%
      split(1:nrow(.)) %>%
      pblapply(function(x) chatLLM(x, intruction_prompt = idealPatientPrompt,
                                   var_1 = x[['study_elegibility_criteria']])) %>%
      do.call(dplyr::bind_rows, .) %>%
      toJSON(pretty = TRUE) %>%
      write(file = 'data/out_04_studies.json')
```

## Resultado

::: {.panel-tabset group="lang"}
### Português
Veja a diferença de tamanho entre os critérios originais e o parágrafo gerado para o primeiro estudo. Os critérios têm `r format(nchar(df$study_elegibility_criteria[1]), big.mark = ".")` caracteres; o parágrafo, `r nchar(df$ideal_patient_portuguese[1])`.

```{r}
#| results: asis
cat("> ", df$ideal_patient_portuguese[1], "\n")
```

```{r}
df %>%
  transmute(study_nct_id, ideal_patient_portuguese = clip(ideal_patient_portuguese, 300)) %>%
  slice(2:4) %>%
  kable()
```

### English
Look at the difference in length between the original criteria and the paragraph generated for the first trial. The criteria have `r format(nchar(df$study_elegibility_criteria[1]), big.mark = ",")` characters; the paragraph, `r nchar(df$ideal_patient_english[1])`.

```{r}
#| results: asis
cat("> ", df$ideal_patient_english[1], "\n")
```

```{r}
df %>%
  transmute(study_nct_id, ideal_patient_english = clip(ideal_patient_english, 300)) %>%
  slice(2:4) %>%
  kable()
```
:::
