---
title: "5. Tipo de patrocinador"
---

```{r}
source("_setup.R")
df <- out("out_05_studies")
```

::: {.panel-tabset group="lang"}
### Português
O ClinicalTrials.gov já informa uma categoria para o patrocinador (`INDUSTRY`, `OTHER`, `NIH`...), mas ela é pouco útil: hospitais, universidades e grupos cooperativos caem todos em `OTHER`. O [`05_ai_sponsor_type.R`](`r gh("05_ai_sponsor_type.R")`){target="_blank"} reclassifica o patrocinador em quatro tipos.

### English
ClinicalTrials.gov already reports a sponsor category (`INDUSTRY`, `OTHER`, `NIH`...), but it isn't very useful: hospitals, universities and cooperative groups all fall under `OTHER`. [`05_ai_sponsor_type.R`](`r gh("05_ai_sponsor_type.R")`){target="_blank"} reclassifies the sponsor into four types.
:::

## O *prompt*

::: {.panel-tabset group="lang"}
### Português
O *prompt* define as quatro categorias com exemplos e dá pistas pelo nome: termos como "Inc.", "Pharmaceuticals" ou "GmbH" indicam indústria; "Group" ou "Network", grupo cooperativo; nomes de hospitais e universidades, instituição de pesquisa. O modelo sempre deve escolher uma categoria, mesmo na dúvida, e devolver também a tradução padronizada em português.

### English
The prompt defines the four categories with examples and gives naming hints: terms like "Inc.", "Pharmaceuticals" or "GmbH" point to industry; "Group" or "Network", to a cooperative group; hospital and university names, to a research facility. The model must always pick a category, even when unsure, and also return the standardized Portuguese translation.
:::

```r
newSponsorClassificationPrompt <- '# Clinical Trial Sponsor Type Classification

## Sponsor Type Categories

- **Cooperative Group**: Multi-institutional research networks/consortia (e.g., NRG Oncology, SWOG, EORTC).
- **Government**: National or governmental health/research bodies (e.g., National Cancer Institute, NIH).
- **Pharmaceutical Industry/Biotech**: For-profit pharmaceutical, biotechnology, or medical device companies.
- **Research Facility**: Individual academic medical centers, universities, hospitals, or cancer centers.

# ... (regras de classificação e tradução)

{{
  "study_ai_lead_sponsor_type": "Cooperative Group | Government | Pharmaceutical Industry/Biotech | Research Facility",
  "study_ai_lead_sponsor_type_portuguese": "Grupo Cooperativo | Governo | Indústria Farmacêutica/Biotecnologia | Instituição de Pesquisa/Hospital"
}}

**NCT ID**: {var_1}
**Trial Title**: {var_2}
**Lead Sponsor Name**: {var_3}'
```

## Aplicando a cada estudo

::: {.panel-tabset group="lang"}
### Português
É a única etapa que usa as três variáveis: o NCT, o título e o nome do patrocinador. O título ajuda quando o nome do patrocinador é ambíguo.

### English
This is the only step that uses all three variables: the NCT ID, the title and the sponsor's name. The title helps when the sponsor's name is ambiguous.
:::

```r
df_studies <- fromJSON('data/out_04_studies.json')

df_studies %>%
      split(1:nrow(.)) %>%
      pblapply(function(x) chatLLM(x, intruction_prompt = newSponsorClassificationPrompt,
                                   var_1 = x[['study_nct_id']],
                                   var_2 = x[['study_brief_title']],
                                   var_3 = x[['study_lead_sponsor_name']])) %>%
      do.call(dplyr::bind_rows, .) %>%
      toJSON(pretty = TRUE) %>%
      write(file = 'data/out_05_studies.json')
```

## Resultado

::: {.panel-tabset group="lang"}
### Português
Repare como os patrocinadores que o ClinicalTrials.gov marca como `OTHER` passam a ter uma categoria clara:

```{r}
show(df, c("study_lead_sponsor_name", "study_lead_sponsor_type", "study_ai_lead_sponsor_type_portuguese"), n = 5)
```

### English
Notice how sponsors that ClinicalTrials.gov marks as `OTHER` now get a clear category:

```{r}
show(df, c("study_lead_sponsor_name", "study_lead_sponsor_type", "study_ai_lead_sponsor_type"), n = 5)
```
:::
