---
title: "1. Downloading the trials"
---

```{r}
source("../../_setup.R")
studies <- out("out_01_studies")
locations <- out("out_01_locations")
```

[`01_clinicaltrials_download.R`](`r gh("01_clinicaltrials_download.R")`){target="_blank"} fetches the trials from the [ClinicalTrials.gov](https://clinicaltrials.gov/data-api/api){target="_blank"} API v2. It's the only step that doesn't use an LLM: it just downloads, tidies and saves.

## How many trials to download

The first line sets how many trials to download. With `5`, the test is quick. With `NULL`, the script downloads every recruiting trial.

```r
n_studies <- 5
```

## Tidying a trial

The API returns each trial as a huge nested list. `tidyClinTrialsStudy()` picks only what matters and turns it into two tables: one row with the trial's information and one row per site.

The helper `collapseField()` handles two common problems: empty fields become `NA`, and fields with more than one value (like phases "PHASE2|PHASE3") become text separated by `|`.

```r
tidyClinTrialsStudy <- function(study){

      protocol <- study$protocolSection
      collapseField <- function(x) if(length(x) == 0) NA_character_ else paste0(unlist(x), collapse = '|')

      studyInfo <- data.frame(
            study_nct_id = protocol$identificationModule$nctId,
            study_acronym = collapseField(protocol$identificationModule$acronym),
            study_brief_title = collapseField(protocol$identificationModule$briefTitle),
            study_official_title = collapseField(protocol$identificationModule$officialTitle),
            study_brief_summary = collapseField(protocol$descriptionModule$briefSummary),
            study_elegibility_criteria = collapseField(protocol$eligibilityModule$eligibilityCriteria),
            study_design_phase = collapseField(protocol$designModule$phases),
            # ... (other columns: description, conditions, status, design, sponsor type)
            study_lead_sponsor_name = collapseField(protocol$sponsorCollaboratorsModule$leadSponsor$name)
      )
```

The sites are in `contactsLocationsModule$locations`. Each one becomes a row, and at the end only the Brazilian sites are kept. International trials often have hundreds of sites worldwide.

```r
      studyLocations <- protocol$contactsLocationsModule$locations %>%
            lapply(function(x) data.frame(
                  study_nct_id = studyInfo$study_nct_id,
                  study_locations.facility = collapseField(x$facility),
                  study_locations.city = collapseField(x$city),
                  study_locations.state = collapseField(x$state),
                  study_locations.zip = collapseField(x$zip),
                  study_locations.country = collapseField(x$country),
                  study_locations.status = collapseField(x$status)
            )) %>%
            dplyr::bind_rows()

      if(nrow(studyLocations) > 0) studyLocations <- studyLocations %>% dplyr::filter(study_locations.country == 'Brazil')

      return(list(study = studyInfo, locations = studyLocations))
}
```

## Querying the API

`getClinicalTrialsStudies()` builds the query with the condition, country and recruitment status. The API returns at most 100 trials per page, plus a `nextPageToken` when there are more. Hence the `while` loop: as long as there's a next page, the script keeps asking. When `n_studies` is a number, it stops after the first page.

```r
getClinicalTrialsStudies <- function(query_condition = 'cancer', location = 'Brazil', status = 'RECRUITING', n_studies = NULL){

      query_search <- list(
            format = 'json',
            query.cond = query_condition,
            query.locn = location,
            filter.overallStatus = status,
            pageSize = ifelse(is.null(n_studies), 100, n_studies)
      )

      search_list <- list()
      has_next_page <- TRUE

      while(has_next_page){

            response <- GET(url = 'https://clinicaltrials.gov/api/v2/studies', query = query_search)
            stop_for_status(response)
            data <- fromJSON(content(response, 'text', encoding = 'UTF-8'), simplifyVector = FALSE)

            search_list <- c(search_list, lapply(data$studies, tidyClinTrialsStudy))
            message(glue('Downloaded Studies: {length(search_list)}'))

            has_next_page <- !is.null(data$nextPageToken) && is.null(n_studies)
            query_search$pageToken <- data$nextPageToken
      }

      return(search_list)
}
```

## Running and saving

Finally, the script calls the function, binds the tables from every trial and saves two files. To build another dataset, for example breast cancer only, just change `query_condition`.

```r
search_list <- getClinicalTrialsStudies(query_condition = 'cancer', location = 'Brazil', status = 'RECRUITING', n_studies = n_studies)

df_studies <- lapply(search_list, function(x) x$study) %>% do.call(dplyr::bind_rows, .)
df_locations <- lapply(search_list, function(x) x$locations) %>% do.call(dplyr::bind_rows, .)

write(toJSON(df_studies, pretty = TRUE), file = 'data/out_01_studies.json')
write(toJSON(df_locations, pretty = TRUE), file = 'data/out_01_locations.json')
```

## Result

`out_01_studies.json` has one row per trial (`r nrow(studies)` in the sample):

```{r}
show(studies, c("study_nct_id", "study_brief_title", "study_design_phase", "study_lead_sponsor_name"), n = 5)
```

And `out_01_locations.json` has one row per trial and site (`r nrow(locations)` in the sample). Notice how many appear only as "Research Site": steps 8 and 9 try to figure out which sites these are.

```{r}
show(locations, c("study_nct_id", "study_locations.facility", "study_locations.city", "study_locations.status"), n = 6)
```
