---
title: "4. Ideal patient"
---

```{r}
source("../../_setup.R")
df <- out("out_04_studies")
```

Eligibility criteria are the hardest part of a clinical trial to read: long inclusion and exclusion lists, often seemingly contradictory at first. [`04_ai_ideal_patient.R`](`r gh("04_ai_ideal_patient.R")`){target="_blank"} turns these criteria into a paragraph describing the patient who would fit the trial.

## The prompt

This is the longest prompt so far. Before writing, the model is told to organize the criteria by topic (diagnosis, prior treatments, performance status...) and cross-check inclusion against exclusion. That way, a criterion like "non-clear cell renal carcinoma" combined with excluding the papillary subtype becomes a single precise sentence.

The final text should have five sentences, start directly with the patient's characteristics and keep the exact medical terms. Lab tests and contraception requirements are left out unless essential.

```r
idealPatientPrompt <- "# Study Patient Profile Generator
## Task
You will receive a clinical studys eligibility criteria and generate a concise, flowing paragraph describing the ideal patient profile for this study in both English and Portuguese.

## Chain-of-Thought Process (Do Not Display in Output)
Before generating the final paragraphs, follow these analytical steps:

1. First, organize all eligibility criteria into categorical groups:
   - Primary diagnosis (disease type, stage, classification)
   - Prior treatments (medications, procedures, therapies)
   - Performance status requirements
   - Exclusion criteria
   # ...
2. Identify and reconcile any complementary or contradictory criteria
3. Prioritize the most clinically relevant aspects
4. Draft and refine 5 comprehensive sentences that capture the essence of the eligible patient

## Requirements for the Patient Profile
# ... (12 rules, see the script)

{{
  \"ideal_patient_english\": \"Your 5-sentence English paragraph here\",
  \"ideal_patient_portuguese\": \"Your 5-sentence Portuguese paragraph here\"
}}

The study eligibility criteria are: {var_1}."
```

## Applying it to each trial

Only one variable is sent here: the full eligibility criteria.

```r
df_studies <- fromJSON('data/out_03_studies.json')

df_studies %>%
      split(1:nrow(.)) %>%
      pblapply(function(x) chatLLM(x, intruction_prompt = idealPatientPrompt,
                                   var_1 = x[['study_elegibility_criteria']])) %>%
      do.call(dplyr::bind_rows, .) %>%
      toJSON(pretty = TRUE) %>%
      write(file = 'data/out_04_studies.json')
```

## Result

Look at the difference in length between the original criteria and the paragraph generated for the first trial. The criteria have `r format(nchar(df$study_elegibility_criteria[1]), big.mark = ",")` characters; the paragraph, `r nchar(df$ideal_patient_english[1])`.

```{r}
#| results: asis
cat("> ", df$ideal_patient_english[1], "\n")
```

```{r}
df %>%
  transmute(study_nct_id, ideal_patient_english = clip(ideal_patient_english, 300)) %>%
  slice(2:4) %>%
  kable()
```
